---
title: "JLC 445/696 Final Project Template"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  cleanrmd::html_document_clean:
    theme: markdown-modest
    toc: yes
    toc_float: 
      collapsed: true
---

Libraries needed for this project (showing this code is optional):
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(zoo)
library(aTSA)
library(tseries)
library(forecast)
```

# 1. Research Question

For example, ***using 2008-2024 data, forecast motor vehicle theft counts and locations in Washington, D.C. for each month in 2025.***

# 2. Data

INSERT ONE PARAGRAPH HERE: getting, cleaning, and filtering the data.

(Showing the code here is optional)

Getting the data:
```{r}
url <- "https://opendata.dc.gov/datasets/DCGIS::crime-incidents-in-"
years <- c(2008:2024)
full.urls <- paste0(url, years, ".csv")
dc.data <- data.frame()

for(file in full.urls) {
  tmp.data <- read.csv(file, stringsAsFactors = FALSE)
  dc.data <- rbind(tmp.data, dc.data)
}
```

Cleaning the data:
```{r}
dc.data <- separate(dc.data, REPORT_DAT, into = c("DATE", "TIME"), sep = " ")
dc.data$DATE <- as.Date(dc.data$DATE, format = "%Y/%m/%d")
dc.data$YEAR <- substr(dc.data$DATE, 0, 4)
dc.data$MONTH <- month(dc.data$DATE)
dc.data$MONTHS <- months(dc.data$DATE)
dc.data$DAY <- day(dc.data$DATE)
dc.data$DOW <- weekdays(dc.data$DATE)
dc.data$HOUR <- substr(dc.data$TIME, 0, 2)
```

Filtering the data:
```{r}
group <- c("MOTOR VEHICLE THEFT")
dc.subset <- dc.data %>% filter(dc.data$OFFENSE %in% group)
```

# 3. Methods

INSERT ONE PARAGRAPH DISCUSSING YOUR ARIMA MODEL AND PARAMETERS (**p**, **d**, and **q**), including: creating crimes per day, making the data stationary, running the ADF test, and running lagged correlations to create your ARIMA model.

(Pick and choose what code you show here; recommend cleaning to only have code for the data you use; meaning don't include 'stationary2' and 'stationary3' if you use 'stationary1')

Create a time series dataset for count of crimes per day:
```{r}
crime <- dc.subset %>%
  group_by(DATE) %>%
  summarise(CRIME.COUNT = n())

crime <- crime %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
crime <- replace(crime, is.na(crime), 0)

cleaned.crime <- zoo(crime$CRIME.COUNT, 
                       seq(from = as.Date(min(crime$DATE)), 
                           to = as.Date(max(crime$DATE)-1), by = 1))
```

Create stationary, differenced datasets:
```{r}
stationary1 <- diff(cleaned.crime, differences = 1)
stationary2 <- diff(cleaned.crime, differences = 2)
stationary3 <- diff(cleaned.crime, differences = 3)
```

Visualize their plots:
```{r}
plot(stationary1)
plot(stationary2)
plot(stationary3)
```

Run ADF tests for the original data, and the differenced data:
```{r}
adf.test(as.matrix(cleaned.crime))
adf.test(as.matrix(stationary1))
adf.test(as.matrix(stationary2))
adf.test(as.matrix(stationary3))
```

Choose the best fitting dataset.

Determine **p** value:
```{r}
pacf(stationary1)
pacf(stationary1, pl=FALSE)
```

Determine the **q** value:
```{r}
acf(stationary1)
acf(stationary1, pl=FALSE)
```

Build an ARIMA model using the parameters identified from above:
```{r}
arima.crime <- auto.arima(cleaned.crime, d = 1, max.p = 10, max.q = 1, seasonal = T)
```

Run the summary stats on your model to identity which parameters were determined to be the best fitting:
```{r}
summary(arima.crime)
```

Check the model residuals:
```{r}
checkresiduals(arima.crime)
```

Forecast every day between now and the end of 2025:
```{r}
forecast.window <- as.numeric(as.Date("2025-12-31")-max(crime$DATE))

forecast.2025 <- forecast(arima.crime, h=forecast.window)
autoplot(forecast.2025)
```

Extract the forecasted values as a table, clean up the column names, add the forecast date and month, filter to just 2025 forecasts, and summarize by month:
```{r}
forecast.values <- as.data.frame(forecast.2025$mean)
forecast.values$ID <- seq.int(nrow(forecast.values))
forecast.upper <- as.data.frame(forecast.2025$upper)
forecast.upper$ID <- seq.int(nrow(forecast.upper))
forecast.values <- forecast.values %>%
  left_join(forecast.upper, by = 'ID')
colnames(forecast.values) <- c("MEAN", "ID", "CI80", "CI90")
forecast.values$DATE <- as.Date(max(crime$DATE) + forecast.values$ID)
forecast.values$MONTH <- months(forecast.values$DATE)

forecast.values.2025 <- subset(forecast.values, forecast.values$DATE > '2024-12-31')
forecast.months <- forecast.values.2025 %>%
  group_by(MONTH) %>%
  summarise(MEAN = round(sum(MEAN),0), FORECAST.90 = round(sum(CI90),0), FORECAST.80 = round(sum(CI80),0))
forecast.months$DIFF <- forecast.months$FORECAST.90 - forecast.months$FORECAST.80
```

# 4. Findings

FIRST PARAGRAPH: describe the model. What do you expect to occur during 2025 - the highs, the lows, the norms, etc. Describe the count of events and likely locations (specific areas of the city). 

SECOND PARAGRAPH: integrate the model. Provide some context to the numbers and locations. Provide at least one reference/citation to a peer-reviewed/academic research study that supports your response.

# 5. Visuals

### 1-2 Graphs

Option 1: Graph actual crime with predicted crime:
```{r}
crime2024 <- crime[c(1,2)]
crime2025 <- forecast.values[c(5,1)]
names(crime2025) <- c("DATE", "CRIME.COUNT")

new.crime <- rbind(crime2024, crime2025)

graph.new.crime <- ggplot(new.crime, aes(x=DATE, y=CRIME.COUNT)) + 
  geom_line() + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Years") + 
  ylab("Crime Count") + 
  ggtitle("TITLE HERE") + 
  geom_area(fill="lightblue", color="black")

graph.new.crime + 
  geom_smooth(method = lm, col = "red", se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Option 2: Graph upper bounds (80% and 90% confidence intervals) by month
```{r}
forecast.months$MONTH <- factor(forecast.months$MONTH, levels = forecast.months$MONTH)
forecast.long <- pivot_longer(forecast.months, cols = c(FORECAST.80, DIFF), 
                          names_to = "Category", values_to = "Value")
forecast.long$Category <- gsub("DIFF", "90% Confidence Interval", forecast.long$Category)
forecast.long$Category <- gsub("FORECAST.80", "80% Confidence Interval", forecast.long$Category)

ggplot(forecast.long, aes(x = MONTH, y = Value, fill = fct_rev(Category))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Value), size = 3, colour = 'white', position = position_stack(vjust = 0.5)) + 
  labs(title = "Washington D.C. 2025 Monthly Forecast Upper Bounds", 
       x = "Month", 
       y = "Crime Count") +
  scale_fill_manual(values = c("90% Confidence Interval" = "blue",
                               "80% Confidence Interval" = "grey"),
                    name = "Forecasts") +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August",
                              "September", "October", "November", "December")) + 
  coord_cartesian(ylim = c(0, 900)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Option 3: Graph mean forecast by month
```{r}
ggplot(forecast.months, aes(x = MONTH, y = MEAN)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) +
  coord_cartesian(ylim = c(450, 550)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Washington D.C. 2025 Mean Monthly Forecast", 
       x = "Month", 
       y = "Crime Count")
```

### 1-2 Maps

  - Get the city outline and roads
```{r, warning=FALSE, message=FALSE, results='hide'}
dc.roads <- roads("DC", "District of Columbia")
dc.roads.major <- dc.roads %>%  filter(RTTYP %in% c("I","S","U"))
dc.outline <- county_subdivisions("DC", "District of Columbia")
```

  - Choose a hotspot method
  - Facet by month
  
```{r, warning=FALSE}
ggplot() +
  geom_sf(data = dc.outline, color = "grey") +
  geom_hex(aes(x = LONGITUDE, y = LATITUDE), data = dc.subset, bins = 16) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data = dc.roads.major, color = "black", alpha = 0.4) +
  theme_classic() +
  facet_wrap(~ MONTH)
```

[THE SOURCE CODE FOR THIS FILE WE CREATED IN CLASS ON 12/4/2024 CAN BE FOUND HERE](https://github.com/matthew-danna/crime-analysis/blob/main/JLC%20445%20696/Final%20Project/JLC%20445%20696%20Final%20Project%202024%2012%2004.Rmd)