---
title: "JLC 445 696 Project 3 Template"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(zoo)
library(aTSA)
library(tseries)
library(forecast)
library(data.table)

url <- "https://opendata.dc.gov/datasets/DCGIS::crime-incidents-in-"
years <- c(2008:2024)
full.urls <- paste0(url, years, ".csv")
dc.data <- data.frame()
for(file in full.urls) {
  tmp.data <- read.csv(file, stringsAsFactors = FALSE)
  dc.data <- rbind(tmp.data, dc.data)
}

dc.data <- separate(dc.data, REPORT_DAT, into = c("DATE", "TIME"), sep = " ")
dc.data$DATE <- as.Date(dc.data$DATE, format = "%Y/%m/%d")
dc.data$YEAR <- substr(dc.data$DATE, 0, 4)
dc.data$MONTH <- month(dc.data$DATE)
dc.data$MONTHS <- months(dc.data$DATE)
dc.data$DAY <- day(dc.data$DATE)
dc.data$DOW <- weekdays(dc.data$DATE)
dc.data$HOUR <- substr(dc.data$TIME, 0, 2)

dc.roads <- roads("DC", "District of Columbia")
dc.roads.major <- dc.roads %>%  filter(RTTYP %in% c("I","S","U"))
dc.outline <- county_subdivisions("DC", "District of Columbia")
```

# Day Shift
```{r, echo=FALSE, message=FALSE, warning=FALSE}
dc1 <- subset(dc.data, dc.data$SHIFT == 'DAY' & 
                dc.data$OFFENSE == 'BURGLARY')
dc.november <- subset(dc1, dc1$MONTH == '11')
crime <- dc1 %>%
  group_by(DATE) %>%
  summarise(COUNT = n())

crime <- crime %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
crime <- replace(crime, is.na(crime), 0)
cleaned.data <- zoo(crime$COUNT, 
                    seq(from = as.Date(min(crime$DATE)), 
                        to = as.Date(max(crime$DATE)-1), by = 1))
stationary.data <- diff(cleaned.data)
arima.data <- auto.arima(stationary.data)

month01 <- as.Date("2024-11-01")
month30 <- as.Date("2024-11-30")
start <- as.numeric(month01 - max(crime$DATE))
end <- as.numeric(month30 - max(crime$DATE))
days <- c(start:end)

# forecasts per day
forecast00 <- forecast(arima.data, h=(days[1]-1))
forecast00 <- round(sum(forecast00$upper[,2]),0)
forecast01 <- forecast(arima.data, h=days[1])
forecast01 <- round(sum(forecast01$upper[,2]),0)
forecast02 <- forecast(arima.data, h=days[2])
forecast02 <- round(sum(forecast02$upper[,2]),0)
forecast03 <- forecast(arima.data, h=days[3])
forecast03 <- round(sum(forecast03$upper[,2]),0)
forecast04 <- forecast(arima.data, h=days[4])
forecast04 <- round(sum(forecast04$upper[,2]),0)
forecast05 <- forecast(arima.data, h=days[5])
forecast05 <- round(sum(forecast05$upper[,2]),0)
forecast06 <- forecast(arima.data, h=days[6])
forecast06 <- round(sum(forecast06$upper[,2]),0)
forecast07 <- forecast(arima.data, h=days[7])
forecast07 <- round(sum(forecast07$upper[,2]),0)
forecast08 <- forecast(arima.data, h=days[8])
forecast08 <- round(sum(forecast08$upper[,2]),0)
forecast09 <- forecast(arima.data, h=days[9])
forecast09 <- round(sum(forecast09$upper[,2]),0)
forecast10 <- forecast(arima.data, h=days[10])
forecast10 <- round(sum(forecast10$upper[,2]),0)
forecast11 <- forecast(arima.data, h=days[11])
forecast11 <- round(sum(forecast11$upper[,2]),0)
forecast12 <- forecast(arima.data, h=days[12])
forecast12 <- round(sum(forecast12$upper[,2]),0)
forecast13 <- forecast(arima.data, h=days[13])
forecast13 <- round(sum(forecast13$upper[,2]),0)
forecast14 <- forecast(arima.data, h=days[14])
forecast14 <- round(sum(forecast14$upper[,2]),0)
forecast15 <- forecast(arima.data, h=days[15])
forecast15 <- round(sum(forecast15$upper[,2]),0)
forecast16 <- forecast(arima.data, h=days[16])
forecast16 <- round(sum(forecast16$upper[,2]),0)
forecast17 <- forecast(arima.data, h=days[17])
forecast17 <- round(sum(forecast17$upper[,2]),0)
forecast18 <- forecast(arima.data, h=days[18])
forecast18 <- round(sum(forecast18$upper[,2]),0)
forecast19 <- forecast(arima.data, h=days[19])
forecast19 <- round(sum(forecast19$upper[,2]),0)
forecast20 <- forecast(arima.data, h=days[20])
forecast20 <- round(sum(forecast20$upper[,2]),0)
forecast21 <- forecast(arima.data, h=days[21])
forecast21 <- round(sum(forecast21$upper[,2]),0)
forecast22 <- forecast(arima.data, h=days[22])
forecast22 <- round(sum(forecast22$upper[,2]),0)
forecast23 <- forecast(arima.data, h=days[23])
forecast23 <- round(sum(forecast23$upper[,2]),0)
forecast24 <- forecast(arima.data, h=days[24])
forecast24 <- round(sum(forecast24$upper[,2]),0)
forecast25 <- forecast(arima.data, h=days[25])
forecast25 <- round(sum(forecast25$upper[,2]),0)
forecast26 <- forecast(arima.data, h=days[26])
forecast26 <- round(sum(forecast26$upper[,2]),0)
forecast27 <- forecast(arima.data, h=days[27])
forecast27 <- round(sum(forecast27$upper[,2]),0)
forecast28 <- forecast(arima.data, h=days[28])
forecast28 <- round(sum(forecast28$upper[,2]),0)
forecast29 <- forecast(arima.data, h=days[29])
forecast29 <- round(sum(forecast29$upper[,2]),0)
forecast30 <- forecast(arima.data, h=days[30])
forecast30 <- round(sum(forecast30$upper[,2]),0)

forecast.list <- c(forecast00, forecast01, forecast02, forecast03, 
                   forecast04, forecast05, forecast06, forecast07, 
                   forecast08, forecast09, forecast10, forecast11, 
                   forecast12, forecast13, forecast14, forecast15, 
                   forecast16, forecast17, forecast18, forecast19, 
                   forecast20, forecast21, forecast22, forecast23, 
                   forecast24, forecast25, forecast26, forecast27, 
                   forecast28, forecast29, forecast30)

forecasts.tmp <- data.frame(c(start-1,days))
colnames(forecasts.tmp) <- "DAYS"
forecasts.tmp$ID <- seq.int(nrow(forecasts.tmp))
forecasts.tmp$DATE <- forecasts.tmp$DAYS + max(crime$DATE)
forecasts.tmp$TOTAL <- forecast.list
forecasts <- subset(forecasts.tmp, forecasts.tmp$ID > 1)
forecasts$ID <- seq.int(nrow(forecasts))
forecasts.tmp <- forecasts.tmp[c(2,4)]
names(forecasts.tmp) <- c("ID", "YESTERDAY")
forecasts <- forecasts %>% left_join(forecasts.tmp, by = "ID")
forecasts$TODAY <- forecasts$TOTAL - forecasts$YESTERDAY
```

### Daytime Burglary Forecasts: November 2024
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(forecasts, aes(DATE, TODAY)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  labs(
    x = "Date",
    y = "Expected Crime Count"
  ) +
  theme(axis.text.x= element_text(angle = 90)) +
  scale_x_continuous("Date", labels = as.character(forecasts$DATE), 
                     breaks = forecasts$DATE)
```

### Daytime Burglary Temporal Topology: Novembers, 2008-2023
```{r, echo=FALSE, message=FALSE, warning=FALSE}
crime.year.day.time <- dc.november %>%
  group_by(YEAR, DOW, HOUR) %>%
  summarise(COUNT = n())

ggplot(crime.year.day.time, aes(HOUR, DOW, fill = COUNT)) +
  geom_tile() + 
  scale_fill_gradient(low = "white", high = "red") +
  theme(legend.position = "none") +
  coord_fixed() +
  scale_y_discrete(limits = c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")) +
  theme(axis.text.x = element_text(angle = 90, size = 5)) + 
  theme(axis.text.y = element_text(size = 5)) + 
  facet_wrap(~ YEAR, nrow = 3)
```

### Daytime Burglary Hotspots: Novembers, 2008-2023
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = dc.outline) +
  geom_hex(aes(x = LONGITUDE, y = LATITUDE), data = dc.november, 
           bins = 15) +
  geom_sf(data = dc.roads.major) +
  scale_fill_continuous(type = "viridis") +
  theme_classic() +
  facet_wrap(~ YEAR, nrow = 3)
```

### 4. Analysis
Sentences here.
- Bullets if you want
* Bullets if you want

# Evening Shift
```{r, echo=FALSE, message=FALSE, warning=FALSE}
dc2 <- subset(dc.data, dc.data$SHIFT == 'EVENING' & 
                dc.data$OFFENSE == 'BURGLARY')
dc.november <- subset(dc2, dc2$MONTH == '11')
crime <- dc2 %>%
  group_by(DATE) %>%
  summarise(COUNT = n())

crime <- crime %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
crime <- replace(crime, is.na(crime), 0)
cleaned.data <- zoo(crime$COUNT, 
                    seq(from = as.Date(min(crime$DATE)), 
                        to = as.Date(max(crime$DATE)-1), by = 1))
stationary.data <- diff(cleaned.data)
arima.data <- auto.arima(stationary.data)

month01 <- as.Date("2024-11-01")
month30 <- as.Date("2024-11-30")
start <- as.numeric(month01 - max(crime$DATE))
end <- as.numeric(month30 - max(crime$DATE))
days <- c(start:end)

# forecasts per day
forecast00 <- forecast(arima.data, h=(days[1]-1))
forecast00 <- round(sum(forecast00$upper[,2]),0)
forecast01 <- forecast(arima.data, h=days[1])
forecast01 <- round(sum(forecast01$upper[,2]),0)
forecast02 <- forecast(arima.data, h=days[2])
forecast02 <- round(sum(forecast02$upper[,2]),0)
forecast03 <- forecast(arima.data, h=days[3])
forecast03 <- round(sum(forecast03$upper[,2]),0)
forecast04 <- forecast(arima.data, h=days[4])
forecast04 <- round(sum(forecast04$upper[,2]),0)
forecast05 <- forecast(arima.data, h=days[5])
forecast05 <- round(sum(forecast05$upper[,2]),0)
forecast06 <- forecast(arima.data, h=days[6])
forecast06 <- round(sum(forecast06$upper[,2]),0)
forecast07 <- forecast(arima.data, h=days[7])
forecast07 <- round(sum(forecast07$upper[,2]),0)
forecast08 <- forecast(arima.data, h=days[8])
forecast08 <- round(sum(forecast08$upper[,2]),0)
forecast09 <- forecast(arima.data, h=days[9])
forecast09 <- round(sum(forecast09$upper[,2]),0)
forecast10 <- forecast(arima.data, h=days[10])
forecast10 <- round(sum(forecast10$upper[,2]),0)
forecast11 <- forecast(arima.data, h=days[11])
forecast11 <- round(sum(forecast11$upper[,2]),0)
forecast12 <- forecast(arima.data, h=days[12])
forecast12 <- round(sum(forecast12$upper[,2]),0)
forecast13 <- forecast(arima.data, h=days[13])
forecast13 <- round(sum(forecast13$upper[,2]),0)
forecast14 <- forecast(arima.data, h=days[14])
forecast14 <- round(sum(forecast14$upper[,2]),0)
forecast15 <- forecast(arima.data, h=days[15])
forecast15 <- round(sum(forecast15$upper[,2]),0)
forecast16 <- forecast(arima.data, h=days[16])
forecast16 <- round(sum(forecast16$upper[,2]),0)
forecast17 <- forecast(arima.data, h=days[17])
forecast17 <- round(sum(forecast17$upper[,2]),0)
forecast18 <- forecast(arima.data, h=days[18])
forecast18 <- round(sum(forecast18$upper[,2]),0)
forecast19 <- forecast(arima.data, h=days[19])
forecast19 <- round(sum(forecast19$upper[,2]),0)
forecast20 <- forecast(arima.data, h=days[20])
forecast20 <- round(sum(forecast20$upper[,2]),0)
forecast21 <- forecast(arima.data, h=days[21])
forecast21 <- round(sum(forecast21$upper[,2]),0)
forecast22 <- forecast(arima.data, h=days[22])
forecast22 <- round(sum(forecast22$upper[,2]),0)
forecast23 <- forecast(arima.data, h=days[23])
forecast23 <- round(sum(forecast23$upper[,2]),0)
forecast24 <- forecast(arima.data, h=days[24])
forecast24 <- round(sum(forecast24$upper[,2]),0)
forecast25 <- forecast(arima.data, h=days[25])
forecast25 <- round(sum(forecast25$upper[,2]),0)
forecast26 <- forecast(arima.data, h=days[26])
forecast26 <- round(sum(forecast26$upper[,2]),0)
forecast27 <- forecast(arima.data, h=days[27])
forecast27 <- round(sum(forecast27$upper[,2]),0)
forecast28 <- forecast(arima.data, h=days[28])
forecast28 <- round(sum(forecast28$upper[,2]),0)
forecast29 <- forecast(arima.data, h=days[29])
forecast29 <- round(sum(forecast29$upper[,2]),0)
forecast30 <- forecast(arima.data, h=days[30])
forecast30 <- round(sum(forecast30$upper[,2]),0)

forecast.list <- c(forecast00, forecast01, forecast02, forecast03, 
                   forecast04, forecast05, forecast06, forecast07, 
                   forecast08, forecast09, forecast10, forecast11, 
                   forecast12, forecast13, forecast14, forecast15, 
                   forecast16, forecast17, forecast18, forecast19, 
                   forecast20, forecast21, forecast22, forecast23, 
                   forecast24, forecast25, forecast26, forecast27, 
                   forecast28, forecast29, forecast30)

forecasts.tmp <- data.frame(c(start-1,days))
colnames(forecasts.tmp) <- "DAYS"
forecasts.tmp$ID <- seq.int(nrow(forecasts.tmp))
forecasts.tmp$DATE <- forecasts.tmp$DAYS + max(crime$DATE)
forecasts.tmp$TOTAL <- forecast.list
forecasts <- subset(forecasts.tmp, forecasts.tmp$ID > 1)
forecasts$ID <- seq.int(nrow(forecasts))
forecasts.tmp <- forecasts.tmp[c(2,4)]
names(forecasts.tmp) <- c("ID", "YESTERDAY")
forecasts <- forecasts %>% left_join(forecasts.tmp, by = "ID")
forecasts$TODAY <- forecasts$TOTAL - forecasts$YESTERDAY
```

### Evening Burglary Forecasts: November 2024
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(forecasts, aes(DATE, TODAY)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  labs(
    x = "Date",
    y = "Expected Crime Count"
  ) +
  theme(axis.text.x= element_text(angle = 90)) +
  scale_x_continuous("Date", labels = as.character(forecasts$DATE), 
                     breaks = forecasts$DATE)
```

### Evening Burglary Temporal Topology: Novembers, 2008-2023
```{r, echo=FALSE, message=FALSE, warning=FALSE}
crime.year.day.time <- dc.november %>%
  group_by(YEAR, DOW, HOUR) %>%
  summarise(COUNT = n())

ggplot(crime.year.day.time, aes(HOUR, DOW, fill = COUNT)) +
  geom_tile() + 
  scale_fill_gradient(low = "white", high = "red") +
  theme(legend.position = "none") +
  coord_fixed() +
  scale_y_discrete(limits = c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")) +
  theme(axis.text.x = element_text(angle = 90, size = 5)) + 
  theme(axis.text.y = element_text(size = 5)) + 
  facet_wrap(~ YEAR, nrow = 3)
```

### Evening Burglary Hotspots: Novembers, 2008-2023
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = dc.outline) +
  geom_hex(aes(x = LONGITUDE, y = LATITUDE), data = dc.november, 
           bins = 15) +
  geom_sf(data = dc.roads.major) +
  scale_fill_continuous(type = "viridis") +
  theme_classic() +
  facet_wrap(~ YEAR, nrow = 3)
```

### 4. Analysis
Sentences here.
- Bullets if you want
* Bullets if you want

# INSERT MIDNIGHT SHIFT STUFF HERE...
