---
title: "JLC 445/696 Project 2 Template"
author: "My Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
## Libraries
```{r, message=FALSE}
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(zoo)
library(aTSA)
library(tseries)
library(forecast)
library(data.table)
```

```{r, echo=FALSE}
dc.data2023.temp <- fread("https://datagate.dc.gov/search/open/crimes?daterange=1-1-2023%20to%20date&details=true&format=csv")
dc.data2022 <- read.csv("https://opendata.arcgis.com/datasets/f9cc541fc8c04106a05a1a4f1e7e813c_4.csv", stringsAsFactors = FALSE)
dc.data2021 <- read.csv("https://opendata.arcgis.com/datasets/619c5bd17ca2411db0689bb0a211783c_3.csv", stringsAsFactors = FALSE)
dc.data2020 <- read.csv("https://opendata.arcgis.com/datasets/f516e0dd7b614b088ad781b0c4002331_2.csv", stringsAsFactors = FALSE)
dc.data2019 <- read.csv("https://opendata.arcgis.com/datasets/f08294e5286141c293e9202fcd3e8b57_1.csv", stringsAsFactors = FALSE)
dc.data2018 <- read.csv("https://opendata.arcgis.com/datasets/38ba41dd74354563bce28a359b59324e_0.csv", stringsAsFactors = FALSE)
dc.data2017 <- read.csv("https://opendata.arcgis.com/datasets/6af5cb8dc38e4bcbac8168b27ee104aa_38.csv", stringsAsFactors = FALSE)
dc.data2016 <- read.csv("https://opendata.arcgis.com/datasets/bda20763840448b58f8383bae800a843_26.csv", stringsAsFactors = FALSE)
dc.data2015 <- read.csv("https://opendata.arcgis.com/datasets/35034fcb3b36499c84c94c069ab1a966_27.csv", stringsAsFactors = FALSE)
dc.data2014 <- read.csv("https://opendata.arcgis.com/datasets/6eaf3e9713de44d3aa103622d51053b5_9.csv", stringsAsFactors = FALSE)
dc.data2013 <- read.csv("https://opendata.arcgis.com/datasets/5fa2e43557f7484d89aac9e1e76158c9_10.csv", stringsAsFactors = FALSE)
dc.data2012 <- read.csv("https://opendata.arcgis.com/datasets/010ac88c55b1409bb67c9270c8fc18b5_11.csv", stringsAsFactors = FALSE)
dc.data2011 <- read.csv("https://opendata.arcgis.com/datasets/9d5485ffae914c5f97047a7dd86e115b_35.csv", stringsAsFactors = FALSE)
dc.data2010 <- read.csv("https://opendata.arcgis.com/datasets/fdacfbdda7654e06a161352247d3a2f0_34.csv", stringsAsFactors = FALSE)
dc.data2009 <- read.csv("https://opendata.arcgis.com/datasets/73cd2f2858714cd1a7e2859f8e6e4de4_33.csv", stringsAsFactors = FALSE)
dc.data2008 <- read.csv("https://opendata.arcgis.com/datasets/180d56a1551c4e76ac2175e63dc0dce9_32.csv", stringsAsFactors = FALSE)

dc.data.temp <- rbind(dc.data2008, dc.data2009, dc.data2010, dc.data2011, dc.data2012, dc.data2013, dc.data2014, dc.data2015, dc.data2016, dc.data2017, dc.data2018, dc.data2019, dc.data2020, dc.data2021, dc.data2022)
dc.data.temp <- separate(dc.data.temp, REPORT_DAT, into = c("DATE", "TIME"), sep = " ")
dc.data.temp <- dc.data.temp %>%
  select(3:9,12:22)
dc.data.temp$DATE <- as.Date(dc.data.temp$DATE, format = "%Y/%m/%d")
dc.data.temp$NEIGHBORHOOD_CLUSTER <- toupper(dc.data.temp$NEIGHBORHOOD_CLUSTER)
dc.data.temp$HOUR <- substr(dc.data.temp$TIME, 0, 2)

dc.data2023.temp2 <- separate(dc.data2023.temp, REPORT_DAT, into = c("DATE", "TIME"), sep = ", ")
dc.data2023.temp3 <- separate(dc.data2023.temp2, TIME, into = c("TIME", "AM.PM"), sep = " ")
dc.data2023.am <- subset(dc.data2023.temp3, dc.data2023.temp3$AM.PM == 'AM')
dc.data2023.am$HOUR <- substr(dc.data2023.am$TIME, 0, 2)
dc.data2023.am$HOUR <- gsub(":", "", dc.data2023.am$HOUR)
dc.data2023.am$HOUR <- paste("0",dc.data2023.am$HOUR, sep = "")
dc.data2023.am$HOUR <- str_sub(dc.data2023.am$HOUR, start = -2)
dc.data2023.am$HOUR <- gsub("12", "00", dc.data2023.am$HOUR)
dc.data2023.pm <- subset(dc.data2023.temp3, dc.data2023.temp3$AM.PM == 'PM')
dc.data2023.pm$HOUR <- substr(dc.data2023.pm$TIME, 0, 2)
dc.data2023.pm$HOUR <- gsub(":", "", dc.data2023.pm$HOUR)
dc.data2023.pm$HOUR <- as.numeric(dc.data2023.pm$HOUR)
dc.data2023.pm$HOUR <- dc.data2023.pm$HOUR + 12
dc.data2023.pm$HOUR <- as.character(dc.data2023.pm$HOUR)
dc.data2023.pm$HOUR <- gsub("24", "12", dc.data2023.pm$HOUR)

dc.data2023 <- rbind(dc.data2023.am, dc.data2023.pm)

dc.data2023 <- subset(dc.data2023, select = c("CCN", "DATE", "TIME", "SHIFT", "METHOD", "OFFENSE", "BLOCK", "WARD", "ANC", "DISTRICT", "PSA", "NEIGHBORHOOD_CLUSTER", "BLOCK_GROUP", "CENSUS_TRACT", "VOTING_PRECINCT", "LATITUDE", "LONGITUDE", "BID", "HOUR"))
dc.data2023$SHIFT <- toupper(dc.data2023$SHIFT)
dc.data2023$METHOD <- toupper(dc.data2023$METHOD)
dc.data2023$OFFENSE <- toupper(dc.data2023$OFFENSE)
dc.data2023$BLOCK <- toupper(dc.data2023$BLOCK)
dc.data2023$NEIGHBORHOOD_CLUSTER <- toupper(dc.data2023$NEIGHBORHOOD_CLUSTER)
dc.data2023$DATE <- as.Date(dc.data2023$DATE, format = "%m/%d/%Y")

dc.data <- rbind(dc.data.temp, dc.data2023)

dc.data$YEAR <- substr(dc.data$DATE, 0, 4)
dc.data$MONTH <- month(dc.data$DATE)
dc.data$DAY <- day(dc.data$DATE)
dc.data$DOW <- weekdays(dc.data$DATE)
```
## Forecast 1: Shift and crime type here
For Forecast 1, my analysis is based on xxxxxx crime during day shift.
```{r}
dc.day <- subset(dc.data, dc.data$SHIFT == 'DAY' & dc.data$OFFENSE == 'ARSON')
```
Here is the code for my daytime forecast model:
```{r}
crime.day <- dc.day %>%
  group_by(DATE) %>%
  summarise(COUNT = n())

crime.day <- crime.day %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
crime.day <- replace(crime.day, is.na(crime.day), 0)

cleaned.day <- zoo(crime.day$COUNT, 
                    seq(from = as.Date(min(crime.day$DATE)), 
                        to = as.Date(max(crime.day$DATE)-1), by = 1))
stationary.day <- diff(cleaned.day)
arima.day <- auto.arima(stationary.day)
```


#### Bar Graph for November 2023 crimes per day
```{r}

```

#### Temporal topology for day and hour, by year
```{r}

```

#### Hotspot map with DC outline and streets, by year
```{r}

```

#### Paragaph of analysis
One paragraph (4-6 sentences) explaining the findings of your forecast, including where and when you expect this crime to occur in D.C. this month (2pts)

## Forecast 2: Shift and crime type here
For Forecast 2, my analysis is based on xxxxxx crime during evening shift.
```{r}
dc.day <- subset(dc.data, dc.data$SHIFT == 'EVENING' & dc.data$OFFENSE == 'ARSON')
```

## Forecast 3: Shift and crime type here
For the third forecast, my analysis is based on xxxxxx crime during midnight shift.
```{r}
dc.day <- subset(dc.data, dc.data$SHIFT == 'MIDNIGHT' & dc.data$OFFENSE == 'ARSON')
```

```{r, echo=FALSE, message=FALSE, results='hide', fig.show='hide'}
# when you set 'echo' to false, you are hiding the block of code 
# when you set 'message' to false, you are hiding the console text that comes after running your code
# when you set 'results' to hide, you are hiding extra processing that occasionally shows up in the console
# when you set fig.show to hide, you are hiding the plot output
```